\begin{abstract}
In this modern world, reducing the gap between local independent musicians and music enthusiasts is crucial for promoting better exposure and fostering music discovery, underscoring the need for a dedicated mobile application. Alunan addresses this need by creating a specialized online community and music discovery platform designed specifically for local independent musicians and music enthusiasts. This platform enables musicians to connect with their audience more effectively while allowing enthusiasts to explore and discover new music within their local scenes. The development of Alunan followed the Mobile Application Development Lifecycle (MADLC), encompassing phases such as identification, design, development, prototyping, and System Usability Scale (SUS) testing. Key features of the app include user profile creation, the ability to share music with post creation, bookmarking favorite musicians, and providing ratings and reviews. Usability was meticulously evaluated, with feedback from 11 evaluators consisting of six musicians and five music enthusiasts providing highly positive results regarding the app's effectiveness and user experience. Future improvements are suggested, including the integration of gamification elements to boost engagement and the addition of song post previews to enhance the discovery process. By addressing the needs of both musicians and enthusiasts, Alunan aims to create a mutually beneficial environment, promoting greater visibility for musicians and a more immersive music discovery experience for users, thus contributing to a richer and more interactive local music ecosystem.
\vspace{1cm}\par
\noindent\textbf{Keywords:} \textit{Local Independent Musicians, Music Enthusiasts, Online Community, Music Discovery, Mobile Application}
\end{abstract}